---
title: "Assignment 6 Solutions: Vector Operations"
---

<span style="color:#9F281A;"> We want to begin to assess the role of distance from schools in determining the education outcomes for Idahoans. We'll use the `landmarks_pnw.csv` and `cejst_pnw.shp` datasets as the basis for this assignment. You'll need to load the csv and convert it to an `sf` object. **We want to compare the percentage of individuals age 25 or over with less than a high school degree (`HSEF` in the cejst dataset) for of counties within 50km of a school (`MTFCC == K2543`) to those that are more than 50km.** </span> 

You'll need to follow many of the same operations in the video example from class. Your assignment is:

<span style="color:#9F281A;">1. Write out the pseudocode for your analysis</span>

>We'll need to do a few things here including load the data, find the tracts within 50km of a school, and then compare the cejst results. Breaking that into pseudocode would look like:

```{r}
#| eval: false
1. Load each dataset
2. Subset to schools
3. Check geometry validity
4. Align CRS
5. Find tracts within 50km
6. Make Maps
```

>Note that my fift step (find tracts within 50km) is a little vague. There are lots of ways I could do this. It might be more helpful to add some specificity here like:

```{r}
#| eval: false
1. Load each dataset
2. Subset to schools
3. Check geometry validity
4. Align CRS
5. Buffer schools by 50km
6. Select tracts within the buffer and attribute
7. Make Maps
```

>There are other ways to do this too (like calculating the distance), but those are likely to be more computationally intensive so I'll leave it at this.

<span style="color:#9F281A;">2. Translate the pseudocode into code chunks and create the necessary code (You'll need to use things like `st_distance`, `st_buffer`, `st_sym_difference`)</span>

>Loading the data should be pretty straightforward for you by now. We use `read_sf` for the shapefile and `read_csv` for the `landmarks.csv`. We then filter the data here so that we aren't working with the entire landmarks dataset.

```{r}
#| echo: true
#| label: loaddata
#| message: false
library(sf)
library(tidyverse, quietly = TRUE)
library(tmap, quietly = TRUE)
cejst.pnw <- read_sf("data/opt/data/2023/assignment06/cejst_pnw.shp")
landmarks.pnw <- read_csv("data/opt/data/2023/assignment06/landmarks_pnw.csv") %>% 
  filter(., MTFCC == "K2543")
```

>We know that one of the datasets are still in long/lat form so we'll need to make it a `sf` object before checking the geometry makes any sense. We'll also assign the crs here by adding it to the `st_as_sf` call. We also need to make sure that there aren't any empty geometries as that will cause problems for mapping later.

```{r}
#| echo: true
#| label: checkvalidity

landmarks.sf <- landmarks.pnw %>% 
  st_as_sf(., coords = c("longitude", "latitude"), crs=4269)
all(st_is_valid(cejst.pnw))
all(st_is_valid(landmarks.sf))

any(st_is_empty(cejst.pnw))
any(st_is_empty(landmarks.sf))
```

>Looks like all the geometries are valid, but there are some empty geometries in the cejst dataset. We will drop those by using `filter` combined with the negation operator (`!`) and `st_is_empty` to return the rows where `st_is_empty` is not equal to `TRUE`. Then we can move forward with making sure the two datasets are aligned by using `st_transform` to change the CRS. We can verify the alignment using a simple call to the `plot` function.

```{r}
#| echo: true
#| label: align
cejst.pnw <- cejst.pnw %>% 
  filter(., !st_is_empty(.))
landmarks.proj <- landmarks.sf %>% 
  st_transform(., crs=st_crs(cejst.pnw))
plot(st_geometry(cejst.pnw))
plot(st_geometry(landmarks.proj), add=TRUE, col="red")
```

> Now it's time to find the tracts that are within 50km of a school. We can do this a few ways. First, we'll use the `st_buffer` approach. We can also calculate the distance matrix between the schools and the tracts using `st_distance`. This adds a little more complexity as we have to then find the values that are greater than 50km. You'll notice that `st_distance` returns a matrix with a row for each school and a column for each tract. This is a little clumsier to deal with, but more precise than a simple buffer.

```{r}
#| echo: true
#| label: finddist

school.buf <- landmarks.proj %>% 
  st_buffer(., dist=50000) 

school.dist <- st_distance(landmarks.proj, cejst.pnw)
dim(school.dist)
```

>Once we have the buffered "footprint" of the school we can use `st_filter` (which filters using topological relations) combined with the `st_covered_by` predicate to find all of the `cejst.pnw` tracts that are covered by the buffer. Notice that if we use the typical `[]` subset we get over 200 more records. This is because the latter takes all tracts with an intersection (rather than using our covered by criteria.). We can alter this to achieve the same result as the `st_filter` by adding the `op=` argument. Using the distance to the points themselves can be a more conservative way of calculating this, but takes a little more work. First we have to get a list of all of the tracts that fall within 50km (using `st_is_within_distance`), then identify which of those list elements are empty (i.e., no schools are within 50km), then set that as an index to subset our cejst data.

```{r}
#| echo: true
#| label: selecttracts

school.tracts.stf <- cejst.pnw %>% 
  st_filter(x =., y = school.buf, .predicate = st_covered_by)
school.tracts.sbst <- cejst.pnw[school.buf,]
school.tracts.sbst2 <- cejst.pnw[school.buf,, op=st_covered_by]

nrow(school.tracts.stf)
nrow(school.tracts.sbst)
nrow(school.tracts.sbst2)

identical(school.tracts.stf, school.tracts.sbst2)

within50 <- st_is_within_distance(cejst.pnw, landmarks.proj, dist=50000, sparse = TRUE)
within50.idx <- lengths(within50) > 0
school.tracts.sbst3 <- cejst.pnw[within50.idx,]
nrow(school.tracts.sbst3)
```

<span style="color:#9F281A;">3. Make a map for both the percentage of individuals with less than a high school degree in counties within 50km and beyond 50km (i.e. make 2 maps)</span>

>We now have the full cejst dataset and a dataset that is subsetted to the tracts within 50km of a school. Plotting the `HSEF` values for the tracts within 50km of a school is easy enough. Just map a layer that contains all of the tracts and set it's color to gray. Then layer the subsetted features on top. Plotting the values of `HSEF` for the tracts beyond 50km is a little trickier (because we haven't created that dataset yet). We can use the index we created in the previous step to do that here. We can also use the `mutate` function to create an indicator variable using our index and then create a "small multiples" style map that plots the two side by side. We'll learn more about "prettying" up these maps in the later parts of the course.

```{r}
#| echo: true
#| label: makemaps
#| cache: true

tm_shape(cejst.pnw) +
  tm_polygons(col="gray") +
  tm_shape(school.tracts.sbst3) +
  tm_fill(col="HSEF")

noschool.tracts <- cejst.pnw[!within50.idx,]
tm_shape(cejst.pnw) +
  tm_polygons(col="gray") +
  tm_shape(noschool.tracts) +
  tm_fill(col="HSEF")

school.combined <- cejst.pnw %>% 
  mutate(., indist = if_else(lengths(within50) > 0, "within50km", "notWithin50km"))


tm_shape(cejst.pnw) +
  tm_polygons(col="gray") +
  tm_shape(school.combined) +
  tm_fill(col="HSEF") +
  tm_facets(by = c("indist"), nrow = 1)
         
```
