---
title: "Spatial Data as Matrices and Rasters"
subtitle: "HES 505 Fall 2022: Session 10"
author: "Matt Williamson"
format: 
  revealjs:
    theme: mytheme.scss
    slide-number: true
    show-slide-number: print
    self-contained: true  
---

# Today's Plan {background="#43464B" background-image="img/slide_10/NEDUS.gif" background-opacity="0.7"}

::: footer
Image Source: USGS
:::

## Objectives

- By the end of today, you should be able to:

  - Describe the raster data model and its representation in R

  - Access the elements that define a raster

  - Build rasters from scratch using matrix operations and `terra`
  
# Defining the raster data model 
::: footer
Image Source: QGIS User's manual
:::

# The Raster Data Model {background="#0033A0"}

## Defining the Raster Data Model

::: columns
:::{.column width="50%"}
::: {style="font-size: 0.7em"}
*   __Vector data__ describe the "exact" locations of features on a landscape (including a Cartesian landscape)

* __Raster data__ represent spatially continuous phenomena (`NA` is possible)

* Depict the alignment of data on a regular lattice (often a square)
  * Operations mimic those for `matrix` objects in `R` 

* Geometry is implicit; the spatial extent and number of rows and columns define the cell size
:::
:::
::: {.column width="50%"}
```{r}
#| fig-height: 6
#| fig-width: 6
library(terra)
filename <- system.file("ex/elev.tif", package="terra")
r <- rast(filename)
plot(r, col=gray(0:1000/1000), main = "Elevation (m)")
```
:::
:::

## Types of Raster Data

::: columns
:::{.column width="50%"}
```{r}
#| fig-height: 6
#| fig-width: 6
library(stars)
x = 1:5
y = 1:4
d = st_dimensions(x = x, y = y, .raster = c("x", "y"))
m = matrix(runif(20),5,4)
r1 = st_as_stars(r = m, dimensions = d)

r = attr(d, "raster")
r$affine = c(0.2, -0.2)
attr(d, "raster") = r
r2 = st_as_stars(r = m, dimensions = d)

r = attr(d, "raster")
r$affine = c(0.1, -0.3)
attr(d, "raster") = r
r3 = st_as_stars(r = m, dimensions = d)

x = c(1, 2, 3.5, 5, 6)
y = c(1, 1.5, 3, 3.5)
d = st_dimensions(x = x, y = y, .raster = c("x", "y"))
r4 = st_as_stars(r = m, dimensions = d)

grd = st_make_grid(cellsize = c(10,10), offset = c(-130,10), n = c(8,5), crs = st_crs(4326))
r5 = st_transform(grd, "+proj=laea +lon_0=-70 +lat_0=35")

par(mfrow = c(2,3), mar = c(0.1, 1, 1.1, 1))
r1 = st_make_grid(cellsize = c(1,1), n = c(5,4), offset = c(0,0))
plot(r1, main = "regular")
plot(st_geometry(st_as_sf(r2)), main = "rotated")
plot(st_geometry(st_as_sf(r3)), main = "sheared")
plot(st_geometry(st_as_sf(r4, as_points = FALSE)), main = "rectilinear")
plot(st_geometry((r5)), main = "curvilinear")
par(mfrow = c(1,1), mar= c(5.1, 4.1, 4.1, 2.1))
```
:::
::: {.column width="50%"}
::: {style="font-size: 0.7em"}
* __Regular__: constant cell size; axes aligned with Easting and Northing

* __Rotated__: constant cell size; axes not aligned with Easting and Northing

* __Sheared__: constant cell size; axes not parallel

* __Rectilinear__: cell size varies along a dimension

* __Curvilinear__: cell size and orientation dependent on the other dimension

:::
:::
:::

## Types of Raster Data

* __Continuous__: numeric data representing a measurement (e.g., elevation, precipitation)

* __Categorical__: integer data representing factors (e.g., land use, land cover)

## Continuous Rasters

::: columns
:::{.column width="50%"}
```{r}
#| echo: true
mintemp <- rast("ftp://ftp.hafro.is/pub/data/rasters/Iceland_minbtemp.tif")
mintemp
```
:::
::: {.column width="50%"}
```{r}
#| fig-height: 6
#| fig-width: 6
plot(mintemp)
```
:::
:::

## Categorical Rasters

::: columns
::: {.column width="50%"}
::: {style="font-size: 0.7em"}
* Requires a classification matrix and coercion to factor

* `levels` allows you to define the categories
:::
```{r}
#| echo: true
# Create classification matrix
cm <- matrix(c(
  -2, 2, 0,
  2, 4, 1,
  4, 10, 2), ncol = 3, byrow = TRUE)

# Create a raster with integers
temp_reclass <- classify(mintemp, cm)
tempcats <- c("cold", "mild", "warm")
levels(temp_reclass) <- tempcats
cats(temp_reclass)
```
:::
::: {.column width="50%"}
```{r}
#| fig-width: 6
#| fig-height: 6
plot(temp_reclass)
```
:::
:::

## Adding Dimensions

* When data are aligned in space and/or time, more efficient to process as 'cubes' or 'stacks'

* Bands of satellite imagery, multiple predictors, spatio-temporal data
```{r}
#| fig-align: center
# (C) 2021, Jonathan Bahlmann, CC-BY-SA
# https://github.com/Open-EO/openeo.org/tree/master/documentation/1.0/datacubes/.scripts
# based on work by Edzer Pebesma, 2019, here: https://gist.github.com/edzer/5f1b0faa3e93073784e01d5a4bb60eca

# plotting runs via a dummy stars object with x, y dimensions (no bands)
# to not be overly dependent on an input image, time steps and bands
# are displayed by replacing the matrix contained in the dummy stars object
# every time something is plotted

# packages, read input ----
set.seed(1331)
suppressPackageStartupMessages(library(colorspace))
suppressPackageStartupMessages(library(scales))

# make color palettes ----
blues <- sequential_hcl(n = 20, h1 = 211, c1 = 80, l1 = 40, l2 = 100, p1 = 2)
greens <- sequential_hcl(n = 20, h1 = 134, c1 = 80, l1 = 40, l2 = 100, p1 = 2)
reds <- sequential_hcl(n = 20, h1 = 360, c1 = 80, l1 = 40, l2 = 100, p1 = 2)
purples <- sequential_hcl(n = 20, h1 = 299, c1 = 80, l1 = 40, l2 = 100, p1 = 2)
greys <- sequential_hcl(n = 20, h1 = 0, c1 = 0, l1 = 40, l2 = 100, p1 = 2)

# matrices from raster ----
# make input matrices from an actual raster image
input <- read_stars("img/slide_10/iceland_delta_cutout_2.tiff") # this raster needs approx 6x7 format
# if the input raster is changed, every image where a pixel value is written as text needs to be checked and corrected accordingly
input <- input[,,,1:4]
warped <- st_warp(input, crs = st_crs(input), cellsize = 200) # warp to approx. 6x7 pixel

# these are only needed for resampling
warped_highres <- st_warp(warped, crs = st_crs(warped), cellsize = 100) # with different input, cellsize must be adapted
# this is a bit of a trick, because 3:4 is different format than 6:7
# when downsampling, the raster of origin isn't so important anyway
warped_lowres <- st_warp(warped_highres[,1:11,,], crs = st_crs(warped), cellsize = 390)
# plot(warped_lowres)
# image(warped[,,,1], text_values = TRUE)

t1 <- floor(matrix(runif(42, -30, 150), ncol = 7)) # create timesteps 2 and 3 randomly
t2 <- floor(matrix(runif(42, -250, 50), ncol = 7))

# create dummy stars object ----
make_dummy_stars <- function(x, y, d1, d2, aff) {
  m = warped_highres[[1]][1:x,1:y,1] # underlying raster doesn't matter because it's just dummy construct
  dim(m) = c(x = x, y = y) # named dim
  dummy = st_as_stars(m)
  attr(dummy, "dimensions")[[1]]$delta = d1
  attr(dummy, "dimensions")[[2]]$delta = d2
  attr(attr(dummy, "dimensions"), "raster")$affine = c(aff, 0.0)
  return(dummy)
}

s <- make_dummy_stars(6, 7, 2.5, -.5714286, -1.14) # mainly used, perspective
f <- make_dummy_stars(6, 7, 1, 1, 0) # flat
highres <- make_dummy_stars(12, 14, 1.25, -.2857143, -.57) # for resampling
lowres <- make_dummy_stars(3, 4, 5, -1, -2) # for resampling

# matrices from image ----
make_matrix <- function(image, band, n = 42, ncol = 7, t = 0) {
  # this is based on an input image with >= 4 input bands
  # n is meant to cut off NAs, ncol is y, t is random matrix for time difference
  return(matrix(image[,,,band][[1]][1:n], ncol = ncol) - t)
  # before: b3 <- matrix(warped[,,,1][[1]][1:42], ncol = 7) - t2
}

# now use function: 
b1 <- make_matrix(warped, 1)
b2 <- make_matrix(warped, 1, t = t1)
b3 <- make_matrix(warped, 1, t = t2)
g1 <- make_matrix(warped, 2)
g2 <- make_matrix(warped, 2, t = t1)
g3 <- make_matrix(warped, 2, t = t2)
r1 <- make_matrix(warped, 3)
r2 <- make_matrix(warped, 3, t = t1)
r3 <- make_matrix(warped, 3, t = t2)
n1 <- make_matrix(warped, 4)
n2 <- make_matrix(warped, 4, t = t1)
n3 <- make_matrix(warped, 4, t = t2)

# plot functions ----
plt = function(x, yoffset = 0, add, li = TRUE, pal, print_geom = TRUE, border = .75, breaks = "equal") {
  # pal is color palette
  attr(x, "dimensions")[[2]]$offset = attr(x, "dimensions")[[2]]$offset + yoffset 
  l = st_as_sf(x, as_points = FALSE)
  if (li)
    pal = lighten(pal, 0.2) # + rnorm(1, 0, 0.1))
  if (! add)
    plot(l, axes = FALSE, breaks = breaks, pal = pal, reset = FALSE, border = grey(border), key.pos = NULL, main = NULL, xlab = "time")
  else
    plot(l, axes = TRUE, breaks = breaks, pal = pal, add = TRUE, border = grey(border))
  u = st_union(l)
  # print(u)
  if(print_geom) {
    plot(st_geometry(u), add = TRUE, col = NA, border = 'black', lwd = 2.5)
  } else {
    # not print geometry
  }
}

pl_stack = function(s, x, y, add = TRUE, nrM, imgY = 7, inner = 1) {
  # nrM is the timestep {1, 2, 3}, cause this function
  # prints all 4 bands at once
  attr(s, "dimensions")[[1]]$offset = x
  attr(s, "dimensions")[[2]]$offset = y
  # m = r[[1]][y + 1:nrow,x + 1:ncol,1]
  m <- eval(parse(text=paste0("n", nrM)))
  s[[1]] = m[,c(imgY:1)] # turn around to have same orientation as flat plot
  plt(s, 0, TRUE,  pal = purples)
  m <- eval(parse(text=paste0("r", nrM)))
  s[[1]] = m[,c(imgY:1)]
  plt(s, 1*inner, TRUE,  pal = reds)
  m <- eval(parse(text=paste0("g", nrM)))
  s[[1]] = m[,c(imgY:1)]
  plt(s, 2*inner, TRUE,  pal = greens)
  m <- eval(parse(text=paste0("b", nrM)))
  s[[1]] = m[,c(imgY:1)]
  plt(s, 3*inner, TRUE, pal = blues) # li FALSE deleted
}

# flat plot function
# prints any dummy stars with any single matrix to position
pl = function(s, x, y, add = TRUE, randomize = FALSE, pal, m, print_geom = TRUE, border = .75, breaks = "equal") {
  # m is matrix to replace image with
  # m <- t(m)
  attr(s, "dimensions")[[1]]$offset = x
  attr(s, "dimensions")[[2]]$offset = y
  # print(m)
  s[[1]] = m
  plt(s, 0, add = TRUE, pal = pal, print_geom = print_geom, border = border, breaks = breaks)
  #plot(s, text_values = TRUE)
}

print_segments <- function(x, y, seg, by = 1, lwd = 4, col = "black") {
  seg = seg * by
  seg[,1] <- seg[,1] + x
  seg[,3] <- seg[,3] + x
  seg[,2] <- seg[,2] + y
  seg[,4] <- seg[,4] + y
  segments(seg[,1], seg[,2], seg[,3], seg[,4], lwd = lwd, col = col)
}

# time series ----

# from: cube1_ts_6x7_bigger.png
offset = 26
plot.new()
#par(mar = c(3, 2, 7, 2))
par(mar = c(0, 0, 0, 0))
#plot.window(xlim = c(10, 50), ylim = c(-3, 10), asp = 1)
plot.window(xlim = c(-15, 75), ylim = c(-3, 10), asp = 1)
pl_stack(s, 0, 0, nrM = 3)
pl_stack(s, offset, 0, nrM = 2)
pl_stack(s, 2 * offset, 0, nrM = 1)
# po <- matrix(c(0,-8,7,0,15,3.5,  0,1,1,5,5,14), ncol = 2)
heads <- matrix(c(3.5, 3.5 + offset, 3.5 + 2*offset, 14,14,14), ncol = 2)
points(heads, pch = 16) # 4 or 16
segments(c(-8, 7, 0, 15), c(-1,-1,3,3), 3.5, 14) # first stack pyramid
segments(c(-8, 7, 0, 15) + offset, c(-1,-1,3,3), 3.5 + offset, 14) # second stack pyramid
segments(c(-8, 7, 0, 15) + 2*offset, c(-1,-1,3,3), 3.5 + 2*offset, 14) # third stack pyramid
arrows(-13, 14, 72, 14, angle = 20, lwd = 2)  # timeline
text(7.5, 3.8, "x", col = "black")
text(-10, -2.5, "bands", srt = 90, col = "black")
text(-4.5, 1.8, "y", srt = 27.5, col = "black")
y = 15.8
text(69, y, "time", col = "black")
text(3.5, y, "2020-10-01", col = "black")
text(3.5 + offset, y, "2020-10-13", col = "black")
text(3.5 + 2*offset, y, "2020-10-25", col = "black")
```
## A note about support
::: {style="font-size: 0.7em"}
* We talked briefly about the `agr` option in the `st_sf()` function

* `agr` refers to the attribute-geometry-relationship which can be:
  * __constant__ = applies to every point in the geometry (lines and polygons are just lots of points)
  * __identity__ = a value unique to a geometry
  * __aggregate__ = a single value that integrates data across the geometry

* __Support__ is the area to which an attribute applies.

* Rasters can have __point__ (attribute refers to the cell center) or __cell__ (attribute refers to an area similar to the pixel) support
:::

# Rasters in `R` {background="#0033A0"}

## Rasters in `R`
::: {style="font-size: 0.7em"}
* `raster` - the original workhorse package; built on `sp`, `rgdal`, and `rgeos`
  * `RasterLayer`, `RasterStack`, and `RasterBrick` classes

* `terra` - relatively new; developed by the `raster` folks, but designed to be much faster
   * `SpatRaster` and `SpatVector` classes

* `stars` - developed by `sf` package developers; `tidyverse` compatible; designed for spatio-temporal data
  * `stars` class
  *  Crosswalk between `raster` and `stars` is available [here](https://github.com/r-spatial/stars/wiki/How-%60raster%60-functions-map-to-%60stars%60-functions)
  * Only way to deal with _rectilinear_ and _curvilinear_ data
:::

## Rasters with `terra`

* syntax is different for `terra` compared to `sf`

* Representation in `Environment` is also different

* Can break pipes, __Be Explicit__

## Rasters by Construction

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
mtx <- matrix(1:16, nrow=4)
mtx
rstr <- terra::rast(mtx)
rstr
```
:::
::: {.column width="50%"}
```{r}
#| fig-width: 6
#| fig-height: 6
plot(rstr)
```
:::
:::
::: {style="font-size: 0.5em"}
Note: you must have `raster` or `terra` loaded for `plot()` to work on `Rast*` objects; otherwise you get `Error in as.double(y) : cannot coerce type 'S4' to vector of type 'double'`
:::

## Rasters by Construction: Origin

* Origin defines the location of the intersection of the x and y axes

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
r <- rast(xmin=-4, xmax = 9.5, ncols=10)
r[] <- runif(ncell(r))
origin(r)
r2 <- r
origin(r2) <- c(2,2) 
```
:::
::: {.column width="50%"}
```{r}
#| fig-width: 6
#| fig-height: 6
par(mfrow = c(2,1))
plot(r, xlim = c(-1, 6), ylim=c(0,7), main = "Original")
plot(r2, xlim = c(-1, 6), ylim=c(0,7), main = "New Origin")
par(mfrow=c(1,1))
```
:::
:::

## Rasters by Construction: Resolution
::: {style="font-size: 0.7em"}
* Geometry is implicit; the spatial extent and number of rows and columns define the cell size
* __Resolution__ (`res`) defines the length and width of an individual pixel
:::

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
r <- rast(xmin=-4, xmax = 9.5, 
          ncols=10)
res(r)
r2 <- rast(xmin=-4, xmax = 5, 
           ncols=10)
res(r2)
```
:::
::: {.column width="50%"}
```{r}
#| echo: true
r <- rast(xmin=-4, xmax = 9.5, 
          res=c(0.5,0.5))
ncol(r)
r2 <- rast(xmin=-4, xmax = 9.5, 
           res=c(5,5))
ncol(r2)
```
:::
:::

## Rasters from Files

* Building rasters useful for templates
* More common to read from files

```{r}
#| echo: true
r <- rast(system.file("ex/elev.tif", package="terra"))
r
```

## Accessing Raster Attributes: Coordinate Reference System

* `terra` stores CRS in WKT format

* Can set and access using `EPSG` and `proj` (deprecated)

* Pay attention to case

```{r}
#| echo: true
r <- rast(system.file("ex/elev.tif", package="terra"))
crs(r)
```

## Accessing Raster Attributes: Coordinate Reference System



```{r}
#| echo: true
r <- rast(system.file("ex/elev.tif", package="terra"))
crs(r, describe=TRUE)
crs(r, proj=TRUE)
crs(r, parse=TRUE)
```

## Accessing Raster Attributes: Bounding box
::: {style="font-size: 0.7em"}
* `terra` uses `ext()` to get or set the extent/bounding box

* Fills cells with `NA`
:::

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
ext(r)
r2 <- r
ext(r2) <- c(5, 7, 48, 52)
ext(r2)
```
:::
::: {.column width="50%"}
```{r}
#| fig-width: 4
#| fig-height: 4
plot(ext(r2))
plot(ext(r), add=TRUE, col="blue", main="Blue is original extent")
```
:::
:::

## Converting vectors to rasters

* Sometimes necessary to convert between data models

* `raster::rasterize`, `terra::rasterize`, `stars::st_rasterize`, and `fasterize::fasterize` all will convert polygons to raster data

* `stars::st_polygonize` will work in the opposite direction

* `terra::vect` will read in vectors as `SpatVectors` or coerce `sf` to `SpatVector`