---
title: "Multivariate Analysis I"
subtitle: "HES 505 Fall 2022: Session 19"
author: "Matt Williamson"
format: 
  revealjs:
    theme: mytheme.scss
    slide-number: true
    show-slide-number: print
    self-contained: true  
---

```{r}
#| include: false
library(sf)
library(tidyverse)
library(maps)
library(tmap)
library(terra)
library(corrplot)
```

# Objectives {background="#0033A0"}

By the end of today you should be able to:

- Recognize the link between regression analysis and overlay analysis

- Generate spatial predictions based on regression analysis

- Extend logistic regression to presence-only data models

## Estimating favorability

$$
\begin{equation}
F(\mathbf{s}) = \prod_{M=1}^{m}X_m(\mathbf{s})
\end{equation}
$$

* Treat $F(\mathbf{s})$ as binary 
* Then $F(\mathbf{s}) = 1$ if all inputs ($X_m(\mathbf{s})$) are suitable
* Then $F(\mathbf{s}) = 0$ if not

## Estimating favorability

$$
\begin{equation}
F(\mathbf{s}) = f(w_1X_1(\mathbf{s}), w_2X_2(\mathbf{s}), w_3X_3(\mathbf{s}), ..., w_mX_m(\mathbf{s}))
\end{equation}
$$

* $F(\mathbf{s})$ does not have to be binary (could be ordinal or continuous)

* $X_m(\mathbf{s})$ could also be extended beyond simply 'suitable/not suitable'

* Adding weights allows incorporation of relative importance

* Other functions for combining inputs ($X_m(\mathbf{s})$)

## Weighted Linear Combinations

$$
\begin{equation}
F(\mathbf{s}) = \frac{\sum_{i=1}^{m}w_iX_i(\mathbf{s})}{\sum_{i=1}^{m}w_i}
\end{equation}
$$

* $F(s)$ is now an index based of the values of $X_m(\mathbf{s})$

* $w_i$ can incorporate weights of evidence, uncertainty, or different participant preferences

* Dividing by $\sum_{i=1}^{m}w_i$ normalizes by the sum of weights

## Model-driven overlay

$$
\begin{equation}
F(\mathbf{s}) = w_0 + \sum_{i=1}^{m}w_iX_i(\mathbf{s}) + \epsilon
\end{equation}
$$

* If we estimate $w_i$ using data, we specify $F(s)$ as the outcome of regression

* When $F(s)$ is binary &rightarrow; logistic regression

* When $F(s)$ is continuous &rightarrow; linear (gamma) regression

* When $F(s)$ is discrete &rightarrow; Poisson regression

* Assumptions about $\epsilon$ matter!!

# Logistic Regression and Distribution Models {background="#0033A0"}

## Why do we create distribution models?

::: columns
::: {.column width="60%"}
::: {style="font-size: 0.7em"}

* To identify important correlations between predictors and the occurrence of an event

* Generate maps of the 'range' or 'niche' of events

* Understand spatial patterns of event co-occurrence

* Forecast changes in event distributions 
:::
:::
::: {.column width="40%"}
![From [Wiens et al. 2009](https://www.pnas.org/content/106/Supplement_2/19729)](img/slide_19/climchange.png)
:::
:::

## General analysis situation 

![From [Long](https://www.biodiversityscience.com/2011/04/27/species-distribution-modelling/)](img/slide_19/SDMfigure1resized.png)

::: {style="font-size: 0.7em"}
* Spatially referenced locations of events $(\mathbf{y})$ sampled from the study extent

* A matrix of predictors $(\mathbf{X})$ that can be assigned to each event based on spatial location

>__Goal__: Estimate the probability of occurrence of events across unsampled regions of the study area based on correlations with predictors

:::

## Modeling Presence-Absence Data

::: columns
::: {.column width="50%"}
* Random or systematic sample of the study region

* The presence (or absence) of the event is recorded for each point

* Hypothesized predictors of occurrence are measured (or extracted) at each point 
:::
::: {.column width="50%"}
![From [By Ragnvald - Own work, CC BY-SA 3.0](https://commons.wikimedia.org/w/index.php?curid=2107716)](img/slide_19/Predicting_habitats.png)
:::
:::

## Logistic regression

::: columns
::: {.column width="50%"}
::: {style="font-size: 0.7em"}

* We can model favorability as the __probability__ of occurrence using a logistic regression

* A _link_ function maps the linear predictor $(\mathbf{x_i}'\beta + \alpha)$ onto the support (0-1) for probabilities

* Estimates of $\beta$ can then be used to generate 'wall-to-wall' spatial predictions
:::
:::
::: {.column width="50%"}

$$ 
\begin{equation}
y_{i} \sim \text{Bern}(p_i)\\
\text{link}(p_i) = \mathbf{x_i}'\beta + \alpha
\end{equation}
$$

![From [Mendoza](https://www.ou.edu/faculty/M/Jorge.L.Mendoza-1/comparison_of_probit.htm)](img/slide_19/Probit.png)

:::
:::

## An Example

Inputs from the `dismo` package

```{r}
base.path <- "/Users/mattwilliamson/Google Drive/My Drive/TEACHING/Intro_Spatial_Data_R/Data/2021/session28/" #sets the path to the root directory

pres.abs <- st_read(paste0(base.path, "presenceabsence.shp"), quiet = TRUE) #read the points with presence absence data
## OGR data source with driver: ESRI Shapefile 
## Source: "/Users/matthewwilliamson/Downloads/session28/presenceabsence.shp", layer: "presenceabsence"
## with 100 features
## It has 1 fields
pred.files <- list.files(base.path,pattern='grd$', full.names=TRUE) #get the bioclim data

pred.stack <- rast(pred.files) #read into a RasterStack
names(pred.stack) <- c("MeanAnnTemp", "TotalPrecip", "PrecipWetQuarter", "PrecipDryQuarter", "MinTempCold", "TempRange")
plot(pred.stack)
```

## An Example

The sample data

::: columns
::: {.column width="40%"}
```{r}
#| echo: true
head(pres.abs)
```
:::
::: {.column width="60%"}
```{r}
#| fig-width: 6
#| fig-height: 6
plot(pred.stack[[1]])
pres.pts <- pres.abs[pres.abs$y == 1,]
abs.pts <- pres.abs[pres.abs$y == 0,]
plot(pres.pts$geometry, add=TRUE, pch=3, col="blue")
plot(abs.pts$geometry, add=TRUE, pch ="-", col="red")
```
:::
:::

## An Example
Building our dataframe


```{r}
#| echo: true
pts.df <- terra::extract(pred.stack, vect(pres.abs), df=TRUE)
head(pts.df)
```

## An Example
Building our dataframe
```{r}
#| echo: true
pts.df[,2:7] <- scale(pts.df[,2:7])
summary(pts.df)
```
## An Example
Looking at correlations
```{r}
#| echo: true
pairs(pts.df[,2:7])
```

## An Example
Looking at correlations

```{r}
#| echo: true
corrplot(cor(pts.df[,2:7]), method = "number")
```

## An Example
Fitting some models

```{r}
#| echo: true
pts.df <- cbind(pts.df, pres.abs$y)
colnames(pts.df)[8] <- "y"
logistic.global <- glm(y~., family=binomial(link="logit"), data=pts.df[,2:8])
logistic.simple <- glm(y ~ MeanAnnTemp + TotalPrecip, family=binomial(link="logit"), data=pts.df[,2:8])
logistic.rich <- glm(y ~ MeanAnnTemp + PrecipWetQuarter + PrecipDryQuarter, family=binomial(link="logit"), data=pts.df[,2:8])
```

## An Example
Checking out the results
```{r}
#| echo: true
summary(logistic.global)
```

## An Example
Checking out the results
```{r}
#| echo: true
summary(logistic.simple)
```

## An Example
Checking out the results
```{r}
#| echo: true
summary(logistic.rich)
```

## An Example
Comparing models
```{r}
#| echo: true
AIC(logistic.global, logistic.simple, logistic.rich)
```

## An Example
Generating predictions

```{r}
#| echo: true
preds <- predict(object=pred.stack, model=logistic.simple)
plot(preds)
plot(pres.pts$geometry, add=TRUE, pch=3, col="blue")
plot(abs.pts$geometry, add=TRUE, pch ="-", col="red")
```

## An Example
Generating predictions

```{r}
#| echo: true
preds <- predict(object=pred.stack, model=logistic.simple, type="response")
plot(preds)
plot(pres.pts$geometry, add=TRUE, pch=3, col="blue")
plot(abs.pts$geometry, add=TRUE, pch ="-", col="red")
```

## An Example
Generating predictions

```{r}
#| echo: true
preds <- predict(object=pred.stack, model=logistic.global, type="response")
plot(preds)
plot(pres.pts$geometry, add=TRUE, pch=3, col="blue")
plot(abs.pts$geometry, add=TRUE, pch ="-", col="red")
```

## An Example
Generating predictions

```{r}
#| echo: true
preds <- predict(object=pred.stack, model=logistic.rich, type="response")
plot(preds)
plot(pres.pts$geometry, add=TRUE, pch=3, col="blue")
plot(abs.pts$geometry, add=TRUE, pch ="-", col="red")
```

## Key assumptions of logistic regression

* Dependent variable must be binary

* Observations must be independent (important for spatial analyses)

* Predictors should not be collinear

* Predictors should be linearly related to the log-odds

* __Sample Size__

# Modelling Presence-Background Data {background="#0033A0"}

## The sampling situation

::: columns
::: {.column width="40%"}
![From [Lentz et al. 2008](https://www.journals.uchicago.edu/doi/full/10.1086/528754)](img/slide_19/maxentresult.png)
:::
::: {.column width="60%"}
* Opportunistic collection of presences only

* Hypothesized predictors of occurrence are measured (or extracted) at each presence

* Background points (or pseudoabsences) generated for comparison
:::
:::

## The Challenge with Background Points

* What constitutes background?

* Not measuring _probability_, but relative likelihood of occurrence

* Sampling bias affects estimation

* The intercept

$$ 
\begin{equation}
y_{i} \sim \text{Bern}(p_i)\\
\text{link}(p_i) = \mathbf{x_i}'\beta + \alpha
\end{equation}
$$

## Point Process Models

* Poisson Point Process Models model location (not $y$)

* Number of points expected is given by a rate $\lambda$

* Model $\lambda$ using Poisson regression

