---
title: "Getting Started"
date: "2023-08-21"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

```

Today we'll focus on getting oriented to the course and the tools we'll be using throughout the semester. Readings are designed to help understand some of the 'rules' of R syntax and develop an understanding for manipulating different types of data in R. I've added a few on open science and reproducibility because I think they help make the case for learning to build code-based workflows.

## Readings

- The [syllabus](/syllabus.qmd), [content](/content/index.qmd), [examples](/example/index.qmd), and [assignments](/assigment/index.qmd) pages for this class

- {{< fa book >}} [Chapter 1 - 6](https://cran.r-project.org/doc/manuals/r-release/R-intro.pdf) in Venables et al., *An Introduction to R* [@venables2009introduction] - for a quick refresher on data types in R (it's only 30 pages)

- {{< fa arrow-up-right-from-square >}} [Chapters 1-2](https://intro2r.com/) in Douglas et al., *An Introduction to R* - provides another intro to R that's been updated and is an open-source book.

- {{< fa arrow-up-right-from-square >}} [Happy Git and GitHub for the useR](https://happygitwithr.com/index.html) - all you really need to know to be a proficient user of git for version control and reproducible workflows.

- {{< fa arrow-up-right-from-square >}} [Open science, reproducibility, and transparency in ecology](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/eap.1822) by Powers and Hampton - discusses the importance of open science for ecologists.

- {{< fa arrow-up-right-from-square >}} [Practical Reproducibility in Geography and Geosciences](https://eartharxiv.org/repository/view/3216/) by Nilst and Pebesma - describes the importance of reproducibility for geospatial analysis.

## Objectives
By the end of today you should:

* Be able to articulate the organization of the course, the approach to grading, and the requirements for the final project

* Be able to access the RStudio Server and Github classroom

* Be able to clone the first self-reflection and assignment and know the process for submitting assignments

## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/01-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/01-slides.html"></iframe>
```

</div>