---
title: "Basic Data Structures in R"
date: "2022-08-24"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "BIhz1dz5Vws", "Basic Data Structures in R", 'title', TRUE,
  "f-sOSXyeeJE", "Accessing subsets of data", "why-not-r", FALSE,
  "Z2foqSMWnGU", "Manipulating and combining data", "Class Details", FALSE,
  "D4LoEctZxHE", "Getting help", "class-details", FALSE
)
```

Today we'll focus on how `R` organizes data and how you can managed and manipulate data in R. Readings are designed to give you more information how `R` conceptualizes data and how various rules and conventions affect your ability to access and manipulate data.

## Readings

- {{< fa arrow-up-right-from-square >}} [Chapter 3](https://intro2r.com/) in Douglas et al., *An Introduction to R* - provides another intro to R that's been updated and is an open-source book.

- {{< fa arrow-up-right-from-square >}} [Writing the perfect question](https://codeblog.jonskeet.uk/2010/08/29/writing-the-perfect-question/) - a blogpost by Jon Skeet that describes the general etiquette for asking coding questions that are likely to get answered.

- {{< fa arrow-up-right-from-square >}} [Reprex Do's and Don'ts](https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html) - provides an overview of the components of a reproducible example and the use of the `reprex` package to generate them.

## Objectives
By the end of today you should:

* Be able to distinguish the major datatypes used in `R` and deploy functions to determine the data type of a given object.

* Be able to subset and manipulate data using base `R ` and the `tidyverse`

* Understand the key elements of getting help with coding errors

## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/02-slides")
slide_tabs(slide_details, "/slides/02-slides.html")
```
