---
title: "Why Geographic Analysis"
date: "2023-08-23"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

```

Before we get inundated with technical details and syntax, I think it's important to remind ourselves why we need geographical analysis. We'll spend some time on the various conceptualizations of place and space, how those things show up in geographic data, and their implications for the kinds of science we can do when we're using geographic data. This session is meant to provide a bit of philosophical foundation for you to keep in mind as you work through various parts of the analytic pipeline.

## Readings



## Objectives
By the end of today you should be able to:

* **Identify** major applications of geographical analysis

* **Describe** critiques and limitations of quantitative geographical analysis 

* **Define** 'scale' and its implications for geographic analysis

* **Place** your research in the broader context of geographic analysis

## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/02-slides")
```

Slides forthcoming