---
title: "Why Geographic Analysis"
date: "2023-08-23"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

```

Before we get inundated with technical details and syntax, I think it's important to remind ourselves why we need geographical analysis. We'll spend some time on the various conceptualizations of place and space, how those things show up in geographic data, and their implications for the kinds of science we can do when we're using geographic data. This session is meant to provide a bit of philosophical foundation for you to keep in mind as you work through various parts of the analytic pipeline.

## Readings

The following readings are intended to give you some sense of the discussion surrounding the role of spatial data in understanding the world. They are a mix of old favorites and relatively recent reviews. You don't need to read all of them or memorize them, but they are worth a skim. I bet you'll find something interesting.

{{< fa arrow-up-right-from-square >}} [Conservation biogeography: assessment and prospects](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1366-9516.2005.00143.x) by Whitaker et al. [-@whittaker2005] provides an overview of the of geography in understanding ecosystem function and shaping conservation strategies.

{{< fa arrow-up-right-from-square >}} [Economic geography, politics, and policy](https://www.annualreviews.org/doi/abs/10.1146/annurev-polisci-050718-033649) by Rickard [-@rickard2020] provides a review of the role of geography in understanding responses to globalization. 

{{< fa arrow-up-right-from-square >}} [Revolutionary and counter revolutionary theory in geography and the problem of ghetto formation](https://onlinelibrary.wiley.com/doi/epdf/10.1111/j.1467-8330.1972.tb00486.x) by David Harvey [-@Harvey1972] offers a scathing critique of quantitative geography (though Harvey was one of the founders of the field). See [@Barnes2009] for a relatively recent attempt to reconcile these views.

{{< fa arrow-up-right-from-square >}} [Does scale exist? An epistemological scale continuum for complex human–environment systems](https://www.sciencedirect.com/science/article/pii/S0016718506001564) by Steven Manson [-@MANSON2008] is one of my favorite summaries of the various definitions and confusion surrounding scale as a concept invoked in many disciplines.

{{< fa arrow-up-right-from-square >}} [Spatial Scaling in Ecology](https://www.jstor.org/stable/2389612) by John Wiens [-@Wiens1989] describes the fundamental challenges of scale in Ecology.


## Objectives
By the end of today you should be able to:

* **Define** what we mean by description, explanation, and prediction in geography.

* **Describe** critiques and limitations of quantitative geographical analysis 

* **Define** 'scale' and its implications for geographic analysis

* **Place** your research in the broader context of geographic analysis

## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/02-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/02-slides.html"></iframe>
```

</div>