---
title: "Introduction to Spatial Data"
date: "2023-08-28"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

```

Now that you have a little more background in the breadth of philosophies, methods, and questions that geography encompasses, it's time to start familiarizing yourself with the nature of spatial data. This lecture will be a little more conceptual, but is designed to help you make some sense for how `R` thinks about spatial data (or at least how the package developers have been thinking about it).

## Readings

{{< fa arrow-up-right-from-square >}} [Types of Spatial Data](https://www.paulamoraga.com/book-spatial/types-of-spatial-data.html) from Paula Moraga's new book Spatial Statistics for Data Science: Theory and Practice with `R` provides a nice overview of the types of spatial data from the perspective of a statistical analyst.

{{< fa arrow-up-right-from-square >}} [Attributes and Support](https://r-spatial.org/book/05-Attributes.html) from Pebesma and Bivand's Spatial Data Science with Applications in `R` gives more info and examples on the nature of the relationship between geometries and support.

{{< fa arrow-up-right-from-square >}} [Scale and Projections](https://open.lib.umn.edu/mapping/chapter/3-scale-and-projections/) from Mapping, Society, and Technology by Laura Matson and Melinda Kernik gives a nice overview of the challenges associated with representing location on Earth's surface.

## Objectives
By the end of today you should be able to:


* **Contrast** the different "views" of spatial data and their incorporation in GIS.

* **Identify** key elements that make data "spatial".

* **Articulate** the importance of coordinate reference systems.

* **Recognize** the relationship between geometries and support.


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/03-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/03-slides.html"></iframe>
```

</div>