---
title: "Repetitive Tasks and Functional Programming"
date: "2022-08-31"
---
```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "BIhz1dz5Vws", "Why use functions?", 'title', TRUE,
  "f-sOSXyeeJE", "Designing functions", "why-not-r", FALSE,
  "Z2foqSMWnGU", "The map and apply families", "Class Details", FALSE,
)
```

## Readings
<i class="fas fa-book"></i> [Functions](https://r4ds.hadley.nz/functions.html) - from Wickham et al., *R for Data Science* [@wickham2016r] describes the logic of functions and how to approach them with `R` and the `tidyverse`.
 
 <i class="fas fa-book"></i> [Iteration](https://r4ds.had.co.nz/iteration.html#iteration) - from Wickham et al., *R for Data Science* [@wickham2016r] provides an introduction to iteration using `for` loops and then builds on that to introduce the  `map` family of functions.

<i class="fas fa-book"></i> [The Apply Family](http://uc-r.github.io/apply_family) - from the UC Business Analytics R Programming Guide provides a nice reference on the `apply` family. These functions are very powerful, but their syntax isn't as intuitive.

## Objectives

By the end of today, you should be able to:

1. Describe the benefit of functions

2. Describe the basic components of functions

3. Design your own simple functions

4. Use the `apply` and `map` families of functions to automate repetitive tasks

## Slides
The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/04-slides")
slide_tabs(slide_details, "/slides/04-slides.html")
```
