---
title: "Spatial Data is Special Data"
date: "2022-09-7"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "BIhz1dz5Vws", "Geography", 'title', TRUE,
  "f-sOSXyeeJE", "Projections", "why-not-r", FALSE,
  "Z2foqSMWnGU", "Spatial Data in R", "Class Details", FALSE,
)
```

## Readings

- <i class="fas fa-book"></i> [Coordinate Reference Systems](https://geocompr.robinlovelace.net/spatial-class.html#crs-intro) Section 2.4 in Lovelace et al., *Geocomputation with R* [@lovelace_geocomputation_2019]

- <i class="fas fa-book"></i> [Chapter 1: Introduction ](https://saylordotorg.github.io/text_essentials-of-geographic-information-systems/s05-introduction.html) in Essentials of Geographic Information Systems by Campbell and Shin [@campbell_essentials_2011]

- <i class="fas fa-external-link-square-alt"></i> [Scale and Projections](https://open.lib.umn.edu/mapping/chapter/3-scale-and-projections/) - a portion of _Mapping, Society, and Technology_ an Laura Matson and Melinda Kernik.

## Objectives
By the end of today, you should be able to:

* Articulate why we care about space

* Describe elements of spatial data

* Define a coordinate reference system and its importance

* Describe several ways to load spatial data into `R`

* Identify, assign, and change projections `R` 


## Slides
The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/06-slides")
slide_tabs(slide_details, "/slides/06-slides.html")
```
