---
title: "Literate Programming, Quarto, and Workflows"
date: "2023-09-6"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))


```

Today we'll focus on some of the tools for reproducible workflows using `R`. We'll introduce Quarto as a means of authoring different kinds of documents. We'll talk about literate programming and leaving breadcrumbs for yourself (and others). Finally, we'll begin to work through the ideas of workflow planning

## Readings

- {{< fa arrow-up-right-from-square >}} [Authoring in Quarto](https://quarto.org/docs/get-started/authoring/rstudio.html) - an intro to Quarto for developing different kinds of documents. Lots of other resources linked here!!

- {{< fa arrow-up-right-from-square >}} [Pseudocode: what it is and how to write it](https://builtin.com/data-science/pseudocode) - A nice blogpost by Sara Metawalli the sketches out the logic of pseudocode and why it can be helpful.

- <i class="fas fa-book"></i> [The Whole Game](https://r4ds.hadley.nz/whole-game.html) - from Wickham et al., *R for Data Science* [@wickham2016r]. Focus on the sections that begin with "Workflow" to get a sense for how we'll start putting the pieces together.

- <i class="fas fa-book"></i> [Scripts, algorithms, and functions](https://geocompr.robinlovelace.net/algorithms.html) - chapter 11 in in Lovelace et al., *Geocomputation with R* [@lovelace_geocomputation_2019] introduces some concepts behind geospatial programming. A few of these pieces will make more sense in the next few weeks, but the general advice on constructing code and planning analyses is useful now.

## Objectives
By the end of today you should be able to:

* **Develop** basic docs with Quarto

* **Understand** the basics of creating readable code

* **Use** pseudocode to sketch out a computational problem

## Slides
The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/06-slides")
```
