---
title: "Areal Data: Rasters"
date: "2023-09-18"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

```

Now that we've learned a bit about how to assess some of the important quantities of vector-based spatial data, we'll try to apply a bit of the same logic to raster data. We'll be using the `terra` package for the majority of raster options in this course primarily because it of its speed. That said, it is not a `tidyverse` package and so some of the intuition we used to organize the `sf` functions will be a little harder to extend here. I'll do my best to help you make the links!!

## Readings

- <i class="fas fa-book"></i> [Chapter 4](https://www.paulamoraga.com/book-spatial/the-terra-package-for-raster-and-vector-data.html) from Paula Moraga's Spatial Statistics for Data Science: Theory and Practice with R provides a quick intro to using `terra` for raster and vector data.

- <i class="fas fa-book"></i> [The `terra` reference page](https://rspatial.github.io/terra/reference/terra-package.html) provides a brief overview of all of the functions and their categories. We'll only focus on the `SpatRaster` methods.

- <i class="fas fa-book"></i> [Raster Data Manipulation](https://rspatial.org/spatial/8-rastermanip.html) from the Spatial Data Science with `R` and `terra` ebook provides some nice examples of `terra` functions in the context of spatial workflows.



## Objectives

By the end of today, you should be able to: 

- Access the elements that define a raster

- Build rasters from scratch using matrix operations and `terra`

- Evaluate logical conditions with raster data

- Calculate different measures of raster data

## Slides
The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/09-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/09-slides.html"></iframe>
```

</div>