---
title: "Building Spatial Databases with Attributes"
date: "2022-10-5"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "tVcomh6jl2Y", "Building spatial databases based on attributes", "title", TRUE,
  "tVcomh6jQ2Y", "Workflows for spatial analysis", "workflows", FALSE,
  "UbQ8IW3UI9E", "Databases and attributes", "database", FALSE,
  "6MTC9uGMNFg", "Common attribute operations", "apps", FALSE,
  "6MTC9uGMNpg", "Joining (a)spatial data", "joins", FALSE,
)
```
## Readings
Today we'll begin exploring typical workflows for spatial analysis by working with attribute data. Attributes generally provide additional information about a location that we can use for visualization and analysis. Unlike spatial operations that we'll explore next week, attribute data do not all require geographic information (but they do need some means of relating to a geography). These chapters are not 'prerequisite' reading for the week, but provide a lot of helpful background for attribute operations in R.

- <i class="fas fa-book"></i> The [Tidy Data](https://r4ds.had.co.nz/tidy-data.html) and [Relational Data](https://r4ds.had.co.nz/relational-data.html) sections from __R For Data Science__ [@wickham2016r] provide a great overview to data cleaning and manipulation functions available in the `tidyverse`.

- <i class="fas fa-book"></i> [Doing things with multiple tables](https://ready4r.netlify.app/labbook/part-5-doing-useful-things-with-multiple-tables.html) has a lot of nice visual examples of for using the `_join` functions in `dplyr`.

- <i class="fas fa-book"></i> This [article](https://www.cell.com/trends/ecology-evolution/fulltext/S0169-5347(21)00227-5) [@di2021conservation] provides a recent recap of a variety of reasons why we may need to combine data from multiple, often disparate, sources. 


## Objectives

By the end of today, you should be able to: 

- Define _spatial analysis_

- Describe the steps in planning a spatial analysis

- Understand the structure of relational databases

- Begin building a database for spatial analysis



## Slides
The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/14-slides")
slide_tabs(slide_details, "/slides/14-slides.html")
```
