---
title: "Operations with Raster Data II"
date: "2023-10-4"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))


```

Now that we've done some "global" transformations of raster data using `terra`, we'll look at some of the options for cell-wise transformations. Rather than manipulating the extent, resolution, or CRS of the raster data; we'll actually be using functions to change the values of the cells themselves.

## Readings
- <i class="fas fa-external-link-square-alt"></i> The [`terra` package vignette](https://rspatial.org/terra/pkg/1-introduction.html#) describes the new raster functions available in `terra`, their relationship to those in the `raster` package, and the changes in syntax between the two.

- <i class="fas fa-book"></i> The [Raster GIS Operations in R with terra](https://www.ecologi.st/spatial-r/raster-gis-operations-in-r-with-terra.html) chapter from Jasper Slingsby's "A Minimal Introduction to GIS (in R)" bookdown project has worked examples of many of the operations we'll learn today.

## Objectives

By the end of today, you should be able to: 

* Use moving windows as a means of smoothing raster data

* Reclassify data using conditional statements and reclassification tables

* Use raster math as a means of creating new data based on an existing dataset.

## Slides
The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/14-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/14-slides.html"></iframe>
```

</div>

[Link to today's Panopto Slides](https://boisestate.hosted.panopto.com/Panopto/Pages/Viewer.aspx?id=d67a2be8-3df8-40a7-a205-b0910143dee4)