---
title: "Building Databases with Location"
date: "2022-10-5"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "tVcomh6jl2Y", "Building spatial databases based on attributes", "title", TRUE,
  "tVcomh6jQ2Y", "Revisting spatial analysis", "workflows", FALSE,
  "UbQ8IW3UI9E", "Estimating spatial attributes", "database", FALSE,
  "6MTC9uGMNFg", "Topological subsetting", "apps", FALSE,
  "6MTC9uGMNpg", "Spatial joins", "joins", FALSE,
)
```

Today we'll continue our development of attributes (or covariates) in our spatial databases. We'll look at developing attributes that describe various geographic properties along with joining and subsetting based on locations.

## Resources
These chapters are not 'prerequisite' reading for the week, but provide a lot of helpful background for determining spatial relations between vector datasets and extracting those into attribute tables for subsequent visualization and analysis.

- <i class="fas fa-book"></i> The [Spatial Data Operations Chapter](https://geocompr.robinlovelace.net/spatial-operations.html) in [@lovelace_geocomputation_2019] makes the concepts of a network concrete (literally) by using a transportation route example to illustrate the various components of a network analysis in `R`.

- <i class="fas fa-book"></i> [Attributes and Support](https://r-spatial.org/book/05-Attributes.html) of Spatial Data Science, a `bookdown` project by Edzer Pebesma and Roger Bivand (of the `sf`, `sp`, `rgeos`, and `rgdal` packages)

## Objectives

By the end of today, you should be able to: 

- Generate new features using geographic data

- Use topological subsetting to reduce features based on geography

- Use spatial joins to add attributes based on location

## Slides
The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/15-slides")
slide_tabs(slide_details, "/slides/15-slides.html")
```