---
title: "Integrating Rasters and Vector Data"
date: "2023-10-16"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

```
The goal of much of our spatial data "munging" is to create a dataframe that can be used in subsequent statistical analyses. It can be difficult to link all of the steps of filtering, selecting, extracting, etc into a coherent problem when you are just being exposed to the syntax (as we discovered last week). Today, I'll try to use a motivating example to help you see a path forward.

## Resources

- <i class="fas fa-book"></i> The [Spatial Data Operations Chapter](https://geocompr.robinlovelace.net/spatial-operations.html) in [@lovelace_geocomputation_2019] makes the concepts of a network concrete (literally) by using a transportation route example to illustrate the various components of a network analysis in `R`.

- <i class="fas fa-book"></i>  [Chapter 3. Processing Tabular Data](https://bookdown.org/mcwimberly/gdswr-book/tables.html) from _Geographic Data Science with R_ by Michael C. Wimberly has a nice introduction to many of the `dplyr` verbs for manipulating tabular data.

- <i class="fas fa-book"></i>  [Chapter 9. Combining Vector Data with Continuous Raster Data](https://bookdown.org/mcwimberly/gdswr-book/combining-vector-data-with-continuous-raster-data.html) from _Geographic Data Science with R_ by Michael C. Wimberly introduces data extraction and zonal statistics for raster data.

- <i class="fas fa-book"></i>  [Chapter 10. Combining Vector Data with Discrete Raster Data](https://bookdown.org/mcwimberly/gdswr-book/tables.html) from _Geographic Data Science with R_ by Michael C. Wimberly extends Chapter 9 for discrete rasters, but also adds some additional buffering and data manipulation syntax.

## Objectives

By the end of today you should be able to:

- Use `dplyr` with `predicates` and `measures` to subset and manipulate data

- Use `extract` to access raster data

- Use `zonal` to summarize access data

- Join data into a single analyzable dataframe


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/16-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/16-slides.html"></iframe>
```

</div>
[Link to Zoom Recording](https://boisestate.zoom.us/rec/share/VWjWFUjEP0QUAw0gbDCfB8AwskdY0qUx_z534eYjuyimlqs1BdQhfYZTfJusTAg2.WMmVMKEYMcahXAG2)