---
title: "Combining Data and Point Patterns"
date: "2023-10-18"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

```

Today we'll finish up our example of combining data for analysis and introduce point process models as a first version of spatial analysis. We'll need a few new packages here, but many of the key data management processes will remain the same.

## Resources

- <i class="fas fa-book"></i> The [Chapters 17 and 18](https://www.paulamoraga.com/book-spatial/spatial-point-patterns.html) on Spatial Point Processes and the `spatstat` package in Paula Moraga's book _Spatial Statistics for Data Science: Theory and Practice with R_.

- <i class="fas fa-external-link-square-alt"></i> [Rings, circles, and null-models for point pattern analysis in ecology](https://onlinelibrary.wiley.com/doi/full/10.1111/j.0030-1299.2004.12497.x?casa_token=D8RfesZ_AjUAAAAA%3AkLmb_B8BubHk4swR3flI9QS4sQnO87wZSImZXsQFD41Q9iEExMEO_bDTPvnvOHtERcit-3YmShhyU0g) by [@wiegand] provides an introduction to metrics for spatial clustering with applications in ecology.

- <i class="fas fa-external-link-square-alt"></i>  [Improving the usability of spatial point process methodology: an interdisciplinary dialogue between statistics and ecology
](https://link.springer.com/article/10.1007/s10182-017-0301-8) by Janine Illian (a major contributor to modern point pattern analyses) and David Burslem (a Scottish plant ecologist) [@illian2017] is a fairly modern take on the challenges associated with point process modeling in ecology.

- <i class="fas fa-external-link-square-alt"></i>  [Chapter 11: Point Pattern Analysis](https://mgimond.github.io/Spatial/chp11_0.html) in Manuel Gimond's Introduction to GIS and Spatial Analysis `bookdown` project provides a nice (and free) introduction to some of these introductory point process methods. 


## Objectives

By the end of today you should be able to:

- Complete the creation of a dataset for analysis using vector and raster data

- Define a point process and their utility for ecological applications

- Define first and second-order Complete Spatial Randomness

- Use several common functions to explore point patterns


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/18-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/18-slides.html"></iframe>
```

</div>

[Link to Panopto Video](https://boisestate.hosted.panopto.com/Panopto/Pages/Viewer.aspx?id=82b8f511-330a-474d-9393-b09f014188de)