---
title: "Interpolation"
date: "2023-10-23"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

```

Point patterns give us the foundation for beginning geostatistical analyses. In geostatistical analyses, we have observations or a spatial process from a limited sample of locations, but would like to be able to infer the values of that process across the entire study region (or at least an area larger than we initially sampled). Interpolation provides one simple way of doing this that relies on the notion that we can learn something about the process simply from our measurements and the location those measurements were taken. We can extend these approaches by adding additional covariates and model structures, but we'll start simple for now.


## Resources

- <i class="fas fa-book"></i> [Chapter 2: Scale](https://boisestatecanvas.instructure.com/files/4856927/download?download_frd=1) in [@fletcher2018spatial] provides a thorough introduction to the ecologist's conceptualization of scale with `R` examples.

- <i class="fas fa-external-link-square-alt"></i>  This [article](https://www.sciencedirect.com/science/article/pii/S0016718506001564#:~:text=The%20epistemological%20scale%20continuum%20runs,readily%20accessible%20to%20objective%20observers) by Steven Manson [@manson2008does] provides a more comprehensive view of conceptualizations of scale.

- <i class="fas fa-external-link-square-alt"></i>  The [Hypothesis Testing and Autocorrelation](https://mgimond.github.io/Spatial/hypothesis-testing.html) chapters of Manuel Gimond's Introduction to GIS and Spatial Analysis `bookdown` project provide concrete examples of attempts to find process from spatial patterns. 

- <i class="fas fa-book"></i> [Chapter 12: Spatial Interpolation](https://r-spatial.org/book/12-Interpolation.html) in Spatial Data Science by Edzer Pebesma and Roger Bivand provides examples of different types of kriging and interpolation using `sf` and `stars`.

## Objectives

By the end of today you should be able to:

*   Distinguish deterministic and stochastic processes

*   Define autocorrelation and describe its estimation

*   Articulate the benefits and drawbacks of autocorrelation

*   Leverage point patterns and autocorrelation to interpolate missing data


```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/19-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/19-slides.html"></iframe>
```

</div>

[Link to Panopto Recording](https://boisestate.hosted.panopto.com/Panopto/Pages/Viewer.aspx?id=9b0b208a-44fc-4c86-bab5-b0a4014222c9)