---
title: "Proximity and Areal Data"
date: "2023-10-25"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))


```

Last class we started to explore ways to leverage spatial autocorrelation as a means of using interpolation to generate values at unobserved locations. We'll continue that discussion using variograms and kriging. We then move to a discussion of areal data and the need to identify "neighbors" as a means of understanding how to weight observations when the actual point location of the observation may be unknown or impossible to assign.


## Resources

- <i class="fas fa-external-link-square-alt"></i> [Ch. 7: Spatial Neighborhood Matrices](https://www.paulamoraga.com/book-spatial/spatial-neighborhood-matrices.html) in
from Paula Moraga's new book Spatial Statistics for Data Science: Theory and Practice with `R` gives a little gentler introduction to spatial neighbors specifically in the context of statistical models.

- <i class="fas fa-book"></i> [Chapter 14 Proximity and Areal Data](https://r-spatial.org/book/14-Areal.html) in Spatial Data Science by Edzer Pebesma and Roger Bivand provides explanations of how the `spdep` package can be used to construct neighborhood weights.

## Objectives

By the end of today you should be able to:

* Describe and implement statistical approaches to interpolation

* Describe the case for identifying neighbors with areal data

* Implement contiguity-based neighborhood detection approaches

* Implement graph-based neighborhood detection approaches



```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/20-slides")
```

<div>

```{=html}
<iframe class="slide-deck" src="/slides/20-slides.html"></iframe>
```

</div>

[Panopto Recording](https://boisestate.hosted.panopto.com/Panopto/Pages/Viewer.aspx?id=976ea29f-b083-4c97-a25d-b0a601435034)
