---
title: "Multivariate Analysis III"
date: "2022-11-02"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "tVcomh6jl2Y", "Roles of Models", "dens", TRUE,
  "tVcomh6jl2Y", "Test Statistics", "dens", TRUE,
  "UbQ8IW3UI9E", "Misclassification", "database", FALSE,
  "6MTC9uGMNFg", "Cross-validation", "apps", FALSE,
)
```

In our last class on multivariate analysis, we'll take on one of the more underappreciated elements of modeling: understanding if your model is good enough for prediction or inference. We'll spend a bit of time differentiating the uses of models as a means of understanding what it means to be a "good" model.

## Resources

- <i class="fas fa-external-link-square-alt"></i> [A practical guide to selecting models for exploration, inference, and prediction in ecology](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.3336) by [@Tredennick2021] highlights the importance of understanding model performance before making inference on predictor effects.

- <i class="fas fa-external-link-square-alt"></i> [Model selection using information criteria, but is the “best” model any good?](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/1365-2664.13060) by [@Mac_Nally2018] highlights the importance of understanding model performance before making inference on predictor effects.

- <i class="fas fa-external-link-square-alt"></i> [Standards for distribution models in biodiversity assessments](https://www.science.org/doi/10.1126/sciadv.aat4858) by [@Araujo2019] highlights the importance of understanding model performance before making inference on predictor effects.

## Objectives

By the end of today you should be able to:

- Articulate three different reasons for modeling and how they link to assessments of fit

- Describe and implement several test statistics for assessing model fit

- Describe and implement several assessments of classification

- Describe and implement resampling techniques to estimate predictive performance

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/21-slides")
slide_tabs(slide_details, "/slides/21-slides.html")
```


