---
title: "Welcome to Space!!"
author: "Matt Williamson"
date: "24 August 2021"
output:
  xaringan::moon_reader:
    lib_dir: "libs"
    chakra: "libs/remark-latest.min.js"
    css: ["default", "css/maw-theme.css", "css/animate.css"]        
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      navigation:
        scroll: false
      # beforeInit: "libs/offline-mathjax.js"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
                      fig.retina = 3, fig.align = "center")
```


```{r packages-data, include=FALSE}
library(tidyverse)
```



class: center middle main-title section-title-4

# Welcome to Space!!

.class-info[

**Session 1**

.light[HES597: Introduction to Spatial Data in R<br>
Boise State University Human-Environment Systems<br>
Fall 2021]

]

---

name: outline
class: title title-inv-5

# Plan for today

--

Introductions

--

Why (not) R?

--

Orientation to the course

--

Parting thoughts


---

layout: false
name: introductions
class: center middle section-title section-title-2 animated fadeIn

# Introductions
---

class: title title-inv-5

# Introductions

A bit about me

--

A bit about you

--

[Course Webpage](https://isdrfall21.spaseslab.com)

--

Class during COVID
---

layout: false
name: why-not-r
class: center middle section-title section-title-2 animated fadeIn

# Why (not) R?
---

class: title title-inv-5

# Why R?

Open Source

--

Integrated analysis pipelines

--

Reproducible workflows 
---

# Why Not R?

Coding can be hard...

--

Memory challenges

--

Speed

--

Decision fatigue
---
layout: false
name: class-details
class: center middle section-title section-title-2 animated fadeIn

# Class Details
---
# The Syllabus

- Expectations (Yours and Mine)
- Course Design
- Advanced Topics

---
# Before next time
- Sign up for github (and send me your username)
- Enter your seat number in my.boisestate.edu
- Verify that you are able to login to the RStudio server with the VPN