---
title: "Spatial data is special data"
author: "Matt Williamson"
date: "31 August 2021"
output:
  xaringan::moon_reader:
    lib_dir: "libs"
    chakra: "libs/remark-latest.min.js"
    css: ["default", "css/mawtheme2.css", "css/animate.css"]        
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      navigation:
        scroll: false
      # beforeInit: "libs/offline-mathjax.js"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
                      fig.retina = 3, fig.align = "center")
```


```{r packages-data, include=FALSE}
library(tidyverse)
library(pander)
```


name:1
class: center middle main-title section-title-4

# Spatial data is special data

#### (Or - how does R keep track of all of this spatial stuff?)

.class-info[

**Session 3**

.light[HES597: Introduction to Spatial Data in R<br>
Boise State University Human-Environment Systems<br>
Fall 2021]

]

---

name: outline
class: title title-inv-5

# Plan for today

--

Location, Location, Location

--

The Coordinate Reference System

--

Other elements that define location in R objects

--

Parting thoughts


---

# Questions we can answer with GIS

 - Questions about geographic location
 
 - Questions about geographic distribution
 
 - Questions about geographic association
 
 - Questions about geographic interaction
 
 - Questions about geographic change

---

# Questions we can answer with GIS

 .primary[- __Questions about geographic location__]
 
 .white[
 - Questions about geographic distribution
 
 - Questions about geographic association
 
 - Questions about geographic interaction
 
 - Questions about geographic change
 ]

---

layout: false
name: projections
class: center middle section-title section-title-2 animated fadeIn

# Location, Location, Location!!
### __(sure, but how do we actually describe where things are?)__
---

class: center middle

# Location
---
class: left middle

# Location

--

## __nominal: (potentially contested) place names__

--

## __absolute: the physical location on the earth's surface__

---