--
title: "Building spatial databases using locations"
author: "Matt Williamson"
date: "05 Oct 2021"
output:
  xaringan::moon_reader:
    lib_dir: "libs"
    chakra: "libs/remark-latest.min.js"
    css: ["default", "css/mawtheme2.css", "css/animate.css"]        
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      navigation:
        scroll: false
      # beforeInit: "libs/offline-mathjax.js"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
                      fig.retina = 3, fig.align = "center")
```


```{r packages-data, include=FALSE}
library(tidyverse)
library(pander)
library(sf)
library(igraph)
library(terra)
library(spData)
```
name: 1
class: center middle main-title section-title-4

# Building spatial databases using locations

.class-info[

**Session 13**

.light[HES597: Introduction to Spatial Data in R<br>
Boise State University Human-Environment Systems<br>
Fall 2021]

]

---