---
title: "Data Structures"
date: "2023-9-1"
---

## Data Types and Structures

### Data Types
Okay, now that we have all of those details out of the way, let's take a look at data structures in `R`. As we discussed,`R` has six basic types of data: numeric, integer, logical, complex, character, and raw. For this class, we won't bother with complex or raw as you are unlikely to encounter them in your introductory spatial explorations.

* __Numeric__ data are numbers that contain a decimal. They can also be whole numbers

* __Integers__ are whole numbers (those numbers without a decimal point). 

* __Logical__ data take on the value of either `TRUE` or `FALSE`. There’s also another special type of logical called `NA` to represent missing values.

* __Character data__ represent string values. You can think of character strings as something like a word (or multiple words). A special type of character string is a factor, which is a string but with additional attributes (like levels or an order). Factors become important in the analyses and visualizations we'll attempt later in the course.

There are a variety of ways to learn more about the structure of different data types:

* `class()` - returns the type of object (high level)
* `typeof()` - returns the type of object (low level)
* `length()` tells you about the length of an object
* `attributes()` - does the object have any metadata

```{r datastructure}
num <- 2.2
class(num)
typeof(num)

y <- 1:10 
y
class(y)
typeof(y)
length(y)

b <- "3"
class(b)
is.numeric(b)

c <- as.numeric(b)
class(c)
```

### Data Structures
You can store information in a variety of ways in `R`. The types we are most likely to encounter this semester are:

* __Vectors__: a collection of elements that are typically `character`, `logical`, `integer`, or `numeric`.

```{r makevects}
#sometimes we'll need to make sequences of numbers to facilitate joins
series <- 1:10
series.2 <- seq(10)
series.3 <- seq(from = 1, to = 10, by = 0.1)
series
series.2
series.3
c(series.2, series.3)

class(series.3)
typeof(series.3)
length(series.3)
```
    
  * Missing Data: R supports missing data in most of the data structures we use, but they can lead to some strange behaviors. Here are a few ways to find missing data:
  
```{r missingdata}
x <- c("a", NA, "c", "d", NA)
is.na(x)
anyNA(x)
```
  
* __Matrices__: are an extension of the numeric or character vectors. They are not a separate type of object but simply an atomic vector with dimensions; the number of rows and columns. As with atomic vectors, the _elements of a matrix must be of the same data_. Matrices are the foundation of rasters, which we'll be discussing frequently throughout the course

```{r matrices}
#matrices are filled columnwise in R
m <- matrix(1:6, nrow = 2, ncol = 3)
dim(m)

x <- 1:3
y <- 10:12

a <- cbind(x, y)
dim(a)
a[3,1]


b <- rbind(x, y)
dim(b)
b[1,3]
```



* __Lists__: Lists essentially act like containers in `R` - they can hold a variety of different data types and structures including more lists. We use lists a lot for functional programming in R where we can apply a function to each element in a list. We'll see this with extracting values from multiple rasters. We can extract elements of lists usin `[]` and `[[]]`

```{r listex}
x <- list(1, "a", TRUE, 1+4i)
x

#adding names
xlist <- list(a = "Waldo", b = 1:10, data = head(mtcars))
xlist

xlist[[1]]
xlist[[3]]
xlist[[3]][1]
xlist[[3]][1,2]
xlist[3][1]

```

* __Data Frames__: data frames resemble that tabular datasets you might be used to in spreadsheet programs and are probably one of the most common types of data in `R`. A data frame is a special type of list where every element has the same length (but can have different types of data). We'll be reading in a number of data frames for this first assignment. 

```{r datframeintro}
dat <- data.frame(id = letters[1:10], x = 1:10, y = 11:20)
dat
is.list(dat)
class(dat)

#lots of ways to look at data in data frames
str(dat) #compact summary of the structure of a dataframe
head(dat) #gives the first 6 rows similar to tail()
dim(dat)
colnames(dat)

## accessing elements of a dataframe
dat[1,3]
dat[["y"]]
dat$y
```

* __Tibbles__: are similar to data frames, but allow for lists _within_ columns. They are designed for use with the `tidyverse` (which we'll explore more in future classes), but the primary reason for introducing them here is because they are the foundation of `sf` objects which we'll use frequently in the weeks to come.

```{r tibble}
library(tidyverse)
dat.tib <- tibble(dat)
is.list(dat.tib)
class(dat.tib)

#lots of ways to look at data in data frames
str(dat.tib) #compact summary of the structure of a dataframe
head(dat.tib) #gives the first 6 rows similar to tail()
dim(dat.tib)
colnames(dat.tib)

## accessing elements of a dataframe
dat.tib[1,3]
dat.tib[["y"]]
dat.tib$y
```

Many of the packages used for spatial operations in `R` rely on special objects (e.g., `sf`, `SpatRasters`) that are combinations of these various elemental data types. That is why we are taking a little time to understand them before jumping into spatial data.